---
title: "MILKYS data from MS Access tables"
author: "DHJ"
date: "2024-13-04"
output: html_document
---

*Started aas a copy of 'Milkys_Data_2016_01_18.Rmd'**
*Folder 'C:\Users\DHJ\OneDrive - NIVA\Documents\seksjon 212\COCO\Analyser\Data_extraction'*

For a quick start, go to chunk 24, `cemp_c2_load`

## R libraries used
```{r}
library(plyr)
library(maps)
library(mapdata)
library(ggplot2)
library(reshape2)
library(dplyr)      # group_by
norway.coast <- map("worldHires", "Norway", col="darkgreen", plot=FALSE)
a <- map("worldHires", plot=FALSE)

# Set wider outout (default = 75)
options(width=90)
```

## Define folders and read data
```{r folders}

# basefolder <- "H:/Documents/seksjon 212/COCO/Analyser/Data_extraction"

datafolder <- "C:/Data/CEMP/2015_Milkys"

#workfolder <- paste0(basefolder,"/Analyser")
#setwd(workfolder)
#setwd(basefolder)

```

```{r file_with_coastline_map}
mapfolder <- "K:\\Kart\\N1000\\"
mapfile <- paste(mapfolder, "norge2.shp", sep="")
```

### Read data **cemp**
(Ca 30 seconds)
```{r data_read, cache=TRUE}
cemp <- read.csv(paste0(datafolder, "/DHJ_Torsk_Table.txt"), 
                 header=TRUE, sep = ";", dec = ",", 
                 na.strings = "", stringsAsFactors=FALSE)
# Fish table
# One line per combination of myear + seqno + rlabo + subno
cemp_f <- read.csv(paste0(datafolder, "/dbo_fish_spe.txt"), 
                 header=TRUE, sep = ";", dec = ",", 
                 na.strings = "", stringsAsFactors=FALSE)
# Measurement table
# Several lines per combination of myear + seqno + rlabo + subno
cemp_m <- read.csv(paste0(datafolder, "/dbo_fish_mas.txt"), 
                 header=TRUE, sep = ";", dec = ",", 
                 na.strings = "", stringsAsFactors=FALSE)

# Tissue table
# One line per combination of myear + seqno + rlabo + subno + tissu
cemp_t <- read.csv(paste0(datafolder, "/dbo_fish_tis.txt"), 
                 header=TRUE, sep = ";", dec = ",", 
                 na.strings = "", stringsAsFactors=FALSE)

# Measurements of contaminant concentrations
# One line per combination of myear + seqno + rlabo + subno + tissu
cemp_c <- read.csv(paste0(datafolder, "/dbo_fish_con.txt"), 
                 header=TRUE, sep = ";", dec = ",", 
                 na.strings = "", stringsAsFactors=FALSE)

# Measurements of biological effects
cemp_be <- read.csv(paste0(datafolder, "/dbo_fish_becon.txt"), 
                 header=TRUE, sep = ";", dec = ",", 
                 na.strings = "", stringsAsFactors=FALSE)

# Find out which values that are non-numeric
# Look at the first sign
first_sign <- substr(cemp$valsnf,1,1)
table(first_sign)

# The ones starting with N
table(cemp$valsnf[first_sign %in% "N"])

# The ones starting with a minus
unique(cemp$valsnf[first_sign %in% "-"])[1:6]
unique(cemp$param[first_sign %in% "-"])

# The ones starting with a space
unique(cemp$valsnf[first_sign %in% " "])[1:6]
unique(cemp$param[first_sign %in% " "])

# Let us set all to numeric (losing only 16 "NA" and 4 "ND")
cemp$valsnf <- as.numeric(cemp$valsnf)
```

```{r biol_effects}
tab <- xtabs(~myear + param, cemp_be)
dim(tab)
(param_be <- colnames(tab))
```


## cemp_c: Row-bind with biological effect tables (cemp_be)  

```{r cemp_c_add_biological_effects}
colnames(cemp_c)
colnames(cemp_be)

cemp_be$method <- NA
cemp_be$analys <- NA
cemp_be$x <- NA

nrow(cemp_c)
nrow(cemp_be)
cemp_c <- rbind(cemp_c, cemp_be)
```
## Fix other tables

### cemp_f: fix/check 

```{r cemp_f_fix}

str(cemp_f)
table(cemp_f$inorb, useNA = "always")

# Delete the one inorb=="1" obs:
cemp_f <- subset(cemp_f, !inorb %in% "1")
nrow(cemp_f)

# Also delete inorb==NA
# cemp_f[is.na(cemp_f$inorb),]
cemp_f <- subset(cemp_f, !is.na(inorb))
nrow(cemp_f)

#
# tiswtx, what is that?
#
first_sign <- substr(cemp_f$tiswtx,1,1)
table(first_sign)
unique(cemp_f$tiswtx[first_sign %in% " "])[1:6]
unique(cemp_f$tiswtx[first_sign %in% "L"])
#test <- as.numeric(sub(",",".", cemp_f$tiswtx))
#cbind(cemp_f$tiswtx[1:100], test[1:100])

# Turn into numeric
cemp_f$tiswtx <- as.numeric(sub(",",".", cemp_f$tiswtx))

#
# shlwtx
#
first_sign <- substr(cemp_f$shlwtx,1,1)
table(first_sign)
unique(cemp_f$shlwtx[first_sign %in% "F"])
unique(cemp_f$shlwtx[first_sign %in% "S"])

# Turn into numeric
cemp_f$shlwtx <- as.numeric(sub(",",".", cemp_f$shlwtx))
```

### cemp_t: fix

```{r cemp_t_fix}
#
# tiswtx in dataset cemp_t
#
first_sign <- substr(cemp_t$tiswtx,1,1)
table(first_sign)
unique(cemp_t$tiswtx[first_sign %in% " "])[1:6]
unique(cemp_t$tiswtx[first_sign %in% "a"])
unique(cemp_t$tiswtx[first_sign %in% "b"])
#test <- as.numeric(sub(",",".", cemp_f$tiswtx))
#cbind(cemp_f$tiswtx[1:100], test[1:100])

# Turn into numeric
cemp_t$tiswtx <- as.numeric(sub(",",".", cemp_t$tiswtx))

#
# drywtx in dataset cemp_t
#
first_sign <- substr(cemp_t$drywtx ,1,1)
table(first_sign)
unique(cemp_t$drywtx [first_sign %in% " "])[1:6]
unique(cemp_t$drywtx [first_sign %in% "b"])
#test <- as.numeric(sub(",",".", cemp_f$tiswtx))
#cbind(cemp_f$tiswtx[1:100], test[1:100])

# Turn into numeric
cemp_t$drywtx <- as.numeric(sub(",",".", cemp_t$drywtx))
```


### cemp_m: get rid of NAs and dublettes  

```{r cemp_m_dublettes}
# Chech for dublettes
tab <- xtabs(~paste(myear, seqno, jmpst, sep="_"), cemp_m)
table(tab)
tab[tab>1]

# Delete NA rows
subset(cemp_m, is.na(myear))
cemp_m <- subset(cemp_m, !is.na(myear))

# Check for dublettes again
tab <- xtabs(~paste(myear, seqno, jmpst, sep="_"), cemp_m)
table(tab)
tab[tab>1]

# check data of dublettes
dublettes <- strsplit(rownames(tab[tab>1]), "_")
for (dub in dublettes)
  print(subset(cemp_m, myear==as.numeric(dub[1]) & seqno==as.numeric(dub[2]) & jmpst==dub[3]))

# delete #2 of dublettes
dim(cemp_m)
for (dub in dublettes){
  #dub <- dublettes[[1]]
  sel <- with(cemp_m, myear==as.numeric(dub[1]) & seqno==as.numeric(dub[2]) & jmpst==dub[3])
  delete_index <- which(sel)[2]
  cemp_m <- cemp_m[-delete_index,]
  }
dim(cemp_m)

# Chech for dublettes of just myear + seqno
tab <- xtabs(~paste(myear, seqno, sep="_"), cemp_m)
table(tab)

# check dublettes
dublettes <- strsplit(rownames(tab[tab>1]), "_")
for (dub in dublettes)
  print(subset(cemp_m, myear==as.numeric(dub[1]) & seqno==as.numeric(dub[2]), 
               select=c(rlabo,myear,seqno,shipc,sdate,jmpst,speci,noins)))

# Chech for dublettes of myear + seqno + rlabo
tab <- xtabs(~paste(myear, seqno, rlabo, sep="_"), cemp_m)
table(tab)
```

## cemp_f2: specimen data  

### cemp_f: merge with cemp_m to add station + species  
```{r cemp_f_add_cemp_m_data}
cemp_f <- merge(cemp_f, cemp_m[,c("myear", "seqno", "rlabo", "jmpst", "speci", "sdate")], 
                by=c("myear", "seqno", "rlabo"), 
                all.x=TRUE, all.y=FALSE)
```

### cemp_f2 = Pick records with species with N > 1000 
```{r cemp_f2}
tab <- table(cemp_f$speci)
tab
(spp <- names(tab[tab>1000]))
# "GADU MOR" "LIMA LIM" "MYTI EDU" "NUCE LAP" "PLAT FLE"
cemp_f2 <- subset(cemp_f, speci %in% spp)
```

### Weight-length plots
```{r weight_length_plot1}
ggplot(cemp_f2, aes(x=lnmea, y=wtmea, color=speci)) + geom_point() + facet_wrap(~inorb)
```

```{r weight_length_plot2}
gg <- ggplot(subset(cemp_f, speci %in% "GADU MOR"), aes(x=lnmea, y=wtmea, color=speci)) + geom_point()
gg + geom_vline(aes(xintercept=130))
```

### Clean cemp_f2 table
```{r cemp_f2_weight_length_clean}
#
# Clean weigth and length
#
plot(wtmea~lnmea, data=subset(cemp_f2, lnmea < 130))
sel <- with(cemp_f2, !is.na(lnmea) & lnmea < 130)
sum(sel)
cemp_f2$lnmea2 <- cemp_f2$lnmea
cemp_f2$lnmea2[sel] <- cemp_f2$lnmea[sel]*10
plot(wtmea~lnmea, data=subset(cemp_f2, lnmea > 130))
points(wtmea~lnmea2, data=subset(cemp_f2, lnmea < 130), col="red")

#
#
#
plot(wtmea~lnmea2, cemp_f2)
plot(wtmea~lnmea2, cemp_f2, ylim = c(0,650))
plot(wtmea~lnmea2, cemp_f2, ylim = c(0,100))

# Try weight in kg
# Quite heavy. Maybe skrei? In this case we keep them so far!
sel <- cemp_f2$wtmea < 40
sum(sel, na.rm=TRUE)
plot(wtmea~lnmea2, cemp_f2)
points(cemp_f2$lnmea2[sel], 1000*cemp_f2$wtmea[sel], col="red")
table(cemp_f2$myear[sel])

#
# A couple of more weight outliters set to NA
#
plot(wtmea~lnmea2, subset(cemp_f2, speci %in% "GADU MOR"), ylim=c(0, 3000))

abline(h=1300, col=4, lty=2)
abline(v=700, col=4, lty=2)
sel <- with(cemp_f2, wtmea < 1300 & lnmea > 700 & speci %in% "GADU MOR")
cemp_f2$wtmea[sel] <- NA

# Be aware of one more outlier (ok weight for length, but just very big)
plot(wtmea~lnmea2, subset(cemp_f2, speci %in% "GADU MOR"))

#
# Result
#
cemp_f2$lnmea <- cemp_f2$lnmea2
plot(wtmea~lnmea, cemp_f2)
```

## Unique key combinations in the tables
```{r tabulate_combinations_of_all_tables}
# Combinations in sample table cemp_m
tab <- xtabs(~myear + seqno + rlabo, cemp_m)
table(tab)
# Combinations in fish table cemp_f2
tab <- xtabs(~myear + seqno + rlabo + subno, cemp_f2)
table(tab)
# Combinations in tissue table cemp_t
tab <- xtabs(~myear + seqno + rlabo + subno + tissu, cemp_t)
table(tab)
# Combinations in concentration table cemp_c
tab <- xtabs(~paste(myear, seqno, rlabo, subno, tissu, param, sep="_"), cemp_c)
table(tab)
# dublettes <- strsplit(rownames(tab[tab>2])[410:420], "_")
# Example
#dub <- strsplit(rownames(tab[tab>2])[420], "_")[[1]]
#subset(cemp_c, myear==as.numeric(dub[1]) & seqno==as.numeric(dub[2]) & rlabo==dub[3] &
#         subno==as.numeric(dub[4]) & tissu==dub[5], select=-c(vflag,qflag,basis,method,analys,unit,x))
```


Find out which valsnf (ie concentration) values that are non-numeric
```{r cemp_c_fix_valsnf}
# Look at the first sign
first_sign <- substr(cemp_c$valsnf,1,1)
table(first_sign)

# The ones starting with N
table(cemp_c$valsnf[first_sign %in% "N"])
# ... are set to NA
cemp_c$valsnf[first_sign %in% "N"] <- NA

# Change to numeric
cemp_c$valsnf <- as.numeric(cemp_c$valsnf)
table(is.na(cemp_c$valsnf))
```

**NOTE:** You may stop here and skip to the `load_data`chunk


## cemp_c: fix  

### fix dublettes 1  

```{r cemp_c_fix_dublettes1}
comb <- with(cemp_c, paste(myear, seqno, rlabo, subno, tissu, param, repno, sep="_"))
tab <- table(comb)
table(tab)
for (i in 1:10){
  sel <- comb == names(tab[tab>1][i])
  print(cemp_c[sel,])
  }
```

### fix dublettes 2  

```{r cemp_c_fix_dublettes2}
#
# Check duplicated data (note: grouping without repno)
#
by_param_cont <- subset(cemp_c, select=c(myear, seqno, rlabo, subno, tissu, param, valsnf)) %>% 
                     group_by(myear, seqno, rlabo, subno, tissu, param)
# We count the number of values
cemp_c2_cont_n <- summarise(by_param_cont, n=n())
# We count the number of distinct values of valsnf
cemp_c2_cont_distinct <- by_param_cont %>% summarise_each(funs(first, last, n_distinct))

# The should have the same number of rows:
dim(cemp_c2_cont_distinct)
dim(cemp_c2_cont_n)

# Add number of values to the 'cemp_c2_cont_distinct' table
cemp_c2_cont_distinct$n <- cemp_c2_cont_n$n 

# Every year has ca 1-2% duplicates 
xtabs(~myear + n, cemp_c2_cont_distinct)
  
# Only 1949, the 1980s, 1996 and 2007 have actually different values for the duplicates
xtabs(~myear + n_distinct, cemp_c2_cont_distinct)

# Check the ones in 2007:
x <- subset(cemp_c2_cont_distinct, myear == 2007 & n_distinct > 1)
dim(x)

# Check the 20 first ones
for (i in 1:20){
  x_example <- x[i, 1:6]
  dupl <- subset(cemp_c, myear %in% x_example$myear &
                   seqno %in% x_example$seqno &
                   rlabo %in% x_example$rlabo &
                   subno %in% x_example$subno &
                   tissu %in% x_example$tissu &
                   param %in% x_example$param)
  print(dupl)
}

```

### cemp_c_clean: create by also including repno (SLOW)

NOTE: two very slow procedures here (7 minutes each)

```{r cemp_c_fix_dublettes3}

#
# Check duplicated data
#
by_param_cont <- subset(cemp_c, select=c(myear, seqno, rlabo, subno, tissu, param, repno, valsnf)) %>% 
                     group_by(myear, seqno, rlabo, subno, tissu, repno, param)
# We count the number of values
cemp_c2_cont_n <- summarise(by_param_cont, n=n())
# We count the number of distinct values of valsnf
cemp_c2_cont_distinct <- by_param_cont %>% summarise_each(funs(first, last, n_distinct))

# The should have the same number of rows:
dim(cemp_c2_cont_distinct)
dim(cemp_c2_cont_n)

# So we can add number of values to the 'cemp_c2_cont_distinct' table
cemp_c2_cont_distinct$n <- cemp_c2_cont_n$n 

# Every year has ca 1-2% duplicates 
xtabs(~myear + n, cemp_c2_cont_distinct)
  
# But there are never different values for the duplicates
xtabs(~myear + n_distinct, cemp_c2_cont_distinct)

# So we just pick the first value of 'valsnf'
cemp_c2_cont <- by_param_cont %>% summarise(valsnf = first(valsnf))

#
# Add inorb, vflag, qflag, unit
#
by_param_char <- subset(cemp_c, select=c(myear, seqno, rlabo, subno, tissu, param, repno, inorb, vflag, qflag, unit)) %>% 
                     group_by(myear, seqno, rlabo, subno, tissu, param, repno)
cemp_c2_char <- by_param_char %>% summarise_each(funs(first, last, n_distinct))
cemp_c2_char_n <- by_param_char %>% summarise(n=n())

if (dim(cemp_c2_char)[1] == dim(cemp_c2_char_n)[1]){
  cemp_c2_char$n <- cemp_c2_char_n$n
} else {
  cat("Something is wrong - cemp_c2_char and cemp_c2_char_n should have the same nukber of rows")
}

head(cemp_c2_char)
# Duplicates
xtabs(~myear + inorb_n_distinct, cemp_c2_char)   # Some
xtabs(~myear + vflag_n_distinct, cemp_c2_char)   # A lot
xtabs(~myear + qflag_n_distinct, cemp_c2_char)   # None
xtabs(~myear + unit_n_distinct, cemp_c2_char)    # Some

xtabs(~inorb_n_distinct + vflag_n_distinct + unit_n_distinct, cemp_c2_char)
# 1. vflag only: 5000 lines
# 2. unit only: 195 lines
# 3. unit + inorb: 257 lines

#
# Check out vflag
# 

# 5291 with different values
xtabs(~ vflag_n_distinct, cemp_c2_char)
# 5291 with different values
xtabs(~ vflag_first + vflag_n_distinct, cemp_c2_char)
xtabs(~ is.na(vflag_first) + vflag_n_distinct, cemp_c2_char)

# Save the doublets of vflag
vflag_doubles <- subset(cemp_c2_char, is.na(vflag_first) & vflag_n_distinct > 1)
dim(vflag_doubles)
head(vflag_doubles)

# Check the 20 first ones
for (i in 1:20 + 5000){
  x_example <- vflag_doubles[i, ]
  dupl <- subset(cemp_c, myear %in% x_example$myear &
                   seqno %in% x_example$seqno &
                   rlabo %in% x_example$rlabo &
                   subno %in% x_example$subno &
                   tissu %in% x_example$tissu &
                   param %in% x_example$param)
  print(dupl)
}

return_doublette_bool <- function(i){
  x_example <- vflag_doubles[i, ]
  with(cemp_c, myear %in% x_example$myear &
                   seqno %in% x_example$seqno &
                   rlabo %in% x_example$rlabo &
                   subno %in% x_example$subno &
                   tissu %in% x_example$tissu &
                   param %in% x_example$param)
  }

load(file = "Milkys_Data_2016_01_18_doublette_vars.RData")  # doublette_no, equal_valsnf, equal_unit, equal_inorb

#
# Set doublette no variable
# SLOW! 7 minutes
#

# UNCOMMENT TO RUN
# doublette_no <- rep(NA, nrow(cemp_c))
# for (i in 1:nrow(vflag_doubles)){
#   doublette_no[return_doublette_bool(i)] <- i
#   }
# UNCOMMENT TO RUN (end)

# Check that there are 2 for each number
length(doublette_no)
sum(!is.na(doublette_no))
sum(!is.na(doublette_no))/2
tab <- table(doublette_no)
table(tab)

#
# Check if valsnf, unit and inorb are the same for all doublettes of vflag
# SLOW! ca 7 minutes
#

# UNCOMMENT TO RUN
# n <- nrow(vflag_doubles)
# equal_valsnf <- rep(NA, n)
# equal_unit <- rep(NA, n)
# equal_inorb <- rep(NA, n)
# for (i in 1:n) {
#   dupl <- cemp_c[doublette_no == i,]
#   equal_valsnf[i] <- dupl[1, "valsnf"]==dupl[2, "valsnf"]
#   equal_unit[i] <- dupl[1, "unit"]==dupl[2, "unit"] | (is.na(dupl[1, "unit"]) & is.na(dupl[2, "unit"]))
#   equal_inorb[i] <- dupl[1, "inorb"]==dupl[2, "inorb"]
# }
# 
# mean(equal_valsnf)
# mean(equal_unit, na.rm=TRUE)
# mean(equal_inorb)
# UNCOMMENT TO RUN (end)

# Pick the ones with vflag = NA for deletion
doublette_delete <- doublette_no
doublette_delete[is.na(cemp_c$vflag)] <- NA

# Check that there is 1 for each number
tab <- table(doublette_delete)
table(tab)

# So we can delete all cases where vflag = NA as there exists a vflag = A for the same combination
cemp_c_clean <- cemp_c[is.na(doublette_delete),]

nrow(cemp_c)
nrow(cemp_c_clean)

# UNCOMMENT TO RUN
# save(doublette_no, equal_valsnf, equal_unit, equal_inorb,
#    file = "Milkys_Data_2016_01_18_doublette_vars.RData")
# UNCOMMENT TO RUN (end)
```


### cemp_c_clean2: new dataset based on cemp_c_clean we put together the data set again  
Tins are removed, since there are some observations which have tins both as individual and pooled samples (I or B)
```{r cemp_c_fix_dublettes4}
## First for categorical vars
by_param_char <- subset(cemp_c_clean, select=c(myear, seqno, rlabo, subno, tissu, param, repno, inorb, vflag, qflag, unit)) %>% 
                     group_by(myear, seqno, rlabo, subno, tissu, param, repno)
cemp_c_char <- by_param_char %>% summarise_each(funs(first, last, n_distinct))
cemp_c_char_n <- by_param_char %>% summarise(n=n())
if (nrow(cemp_c_char) == nrow(cemp_c_char_n)){
  cemp_c_char$n <- cemp_c_char_n$n
} else {
  cat("WARNING - Unequal no of rows!")
}

head(cemp_c_char)
# Duplicates
xtabs(~inorb_n_distinct, cemp_c_char)   
xtabs(~vflag_n_distinct, cemp_c_char)
xtabs(~qflag_n_distinct, cemp_c_char)  
xtabs(~unit_n_distinct, cemp_c_char)  
xtabs(~inorb_n_distinct+ unit_n_distinct, cemp_c_char)  
# Also check NA values
xtabs(~inorb_n_distinct+ unit_n_distinct + is.na(inorb_first), cemp_c_char)  
xtabs(~inorb_n_distinct+ unit_n_distinct + is.na(unit_first), cemp_c_char)  

# Check years - only 2006-2009
xtabs(~myear + unit_n_distinct, cemp_c_char)  

# Check parameters - only Tins (DBTIN, DPTIN, MBTIN, MPTIN, TPTIN) 
tab <- xtabs(~param + unit_n_distinct, cemp_c_char)
tab[tab[,2]>0,]
par_tins <- rownames(tab)[tab[,2]>0]

#
# Have a look at 452 unit doublettes
#   of which 257 also are inorb doublettes
#

# Save the doublets of unit 
unit_doubles <- subset(cemp_c_char,  unit_n_distinct > 1)
dim(unit_doubles)
head(unit_doubles)

return_doublette_bool <- function(i){
  x_example <- unit_doubles[i, ]
  with(cemp_c_clean, myear %in% x_example$myear &
                   seqno %in% x_example$seqno &
                   rlabo %in% x_example$rlabo &
                   subno %in% x_example$subno &
                   tissu %in% x_example$tissu &
                   param %in% x_example$param)
  }

# check some examples, and check those in the fish table
# Some are dog whelks (NUCE LAP), some are cod (GADU MOR)

random_unit_doubles <- sample(1:nrow(unit_doubles), 20)
for (i in random_unit_doubles){
  x_example <- cemp_c_clean[return_doublette_bool(i),]
  sel <- with(cemp_f2, myear %in% x_example$myear &
                   seqno %in% x_example$seqno &
                   rlabo %in% x_example$rlabo &
                   subno %in% x_example$subno)
  print(x_example)
  print(cemp_f2[sel, "speci"])
  }

#
# For now, we just delete all the tins
#

cemp_c_clean2 <- subset(cemp_c_clean, !param %in% par_tins)
nrow(cemp_c_clean)
nrow(cemp_c_clean2)
```


### cemp_c_clean2: Final check  
Based on cemp_c_clean2 we put together the data set again
```{r cemp_c_fix_dublettes5}
#
# First for categorical vars
#
by_param_char <- subset(cemp_c_clean2, select=c(myear, seqno, rlabo, subno, tissu, param, repno, inorb, vflag, qflag, unit)) %>% 
                     group_by(myear, seqno, rlabo, subno, tissu, param, repno)
cemp_c_char <- by_param_char %>% summarise_each(funs(first, last, n_distinct))
cemp_c_char_n <- by_param_char %>% summarise(n=n())
if (nrow(cemp_c_char) == nrow(cemp_c_char_n)){
  cemp_c_char$n <- cemp_c_char_n$n
} else {
  cat("WARNING - Unequal no of rows!")
}

head(cemp_c_char)
# Duplicates
xtabs(~inorb_n_distinct, cemp_c_char)   
xtabs(~vflag_n_distinct, cemp_c_char)
xtabs(~qflag_n_distinct, cemp_c_char)  
xtabs(~unit_n_distinct, cemp_c_char)  
# Only 1's!

#
# Then for the continuous vars
#
by_param_cont <- subset(cemp_c_clean2, select=c(myear, seqno, rlabo, subno, tissu, param, repno, valsnf)) %>% 
                     group_by(myear, seqno, rlabo, subno, tissu, repno, param)
cemp_c_cont_distinct <- by_param_cont %>% summarise_each(funs(first, last, n_distinct))
# We count the number of values
cemp_c_cont_n <- summarise(by_param_cont, n=n())


# They should have the same number of rows:
# So we can add number of values to the 'cemp_c_cont_distinct' table
if (nrow(cemp_c_cont_distinct) == nrow(cemp_c_cont_n)){
  cemp_c_cont_distinct$n <- cemp_c_cont_n$n 
} else {
  cat("WARNING - Unequal no of rows!")
}

# No duplicates... 
xtabs(~n, cemp_c_cont_distinct)
# ... and then of course always one distinct value
xtabs(~n_distinct, cemp_c_cont_distinct)
# again - yeah!
```

### cemp_c2: Remake   
Combine categorical variables (inorb, vflag, qflag, unit) with continuous (valsnf).  
This is the final "concentration" data file (used to )
```{r cemp_c2_create}
if (nrow(cemp_c_char) == nrow(cemp_c_cont_distinct)){
  cemp_c2 <- cbind(cemp_c_char, cemp_c_cont_distinct[,"first"])
} else {
  cat("WARNING - Unequal no of rows!")
  cemp_c2 <- NULL
}
cemp_c2 <- rename(cemp_c2, valsnf = first)
cemp_c2 <- rename(cemp_c2, unit = unit_first, inorb = inorb_first, vflag=vflag_first, qflag=qflag_first)
columns_to_remove <- colnames(cemp_c2) %in% c("inorb_last","vflag_last", "qflag_last", "unit_last",
                                                    "inorb_n_distinct","vflag_n_distinct", 
                                                    "qflag_n_distinct", "unit_n_distinct",
                                                    "n")
which(columns_to_remove)
cemp_c2 <- cemp_c2[,!columns_to_remove]
dim(cemp_c2)
```

### Save 
Saved as R file
```{r cemp_c2_save}
save(cemp_c2, file="Milkys_Data_2016_01_18_cemp_c2.RData")
save(cemp_c2, cemp_t, cemp_f2, cemp_m, file="Milkys_Data_2016_01_18_cemp_tables.RData")
```

Load all R files made/changed so far
```{r cemp_c2_load}
# load(file="Milkys_Data_2016_01_18_cemp_c2.RData")                # cemp_c2
load(file="Milkys_Data_2016_01_18_cemp_tables.RData")            # cemp_c2, cemp_t, cemp_f2, cemp_m
```

## Combine

- Fish table (cemp_f2): One line per combination of myear + seqno + rlabo + subno
- Measurement table: Several lines per combination of myear + seqno + rlabo + subno
- Tissue table (cemp_t): One line per combination of myear + seqno + rlabo + subno + tissu
```{r cemp_c2_to_cemp_comb_leftjoin_with_tissue1}
# tab <- xtabs(~myear + seqno + rlabo, cemp_m)
# table(tab)
# tab <- xtabs(~myear + seqno + rlabo + subno, cemp_f2)
# table(tab)
# tab <- xtabs(~myear + seqno + rlabo + subno + tissu, cemp_t)
# table(tab)

####################################################################################################
#
# Combine with tissue data
#
####################################################################################################

#
# Parameters
#
apply(!is.na(cemp_t), 2, sum)
```

### Comments of cemp_t
Make liver colour variables
```{r cemp_c2_to_cemp_comb_leftjoin_with_tissue2_comment_liver_colour}
# comment 1
tab <- table(cemp_t$cmnt1)
tab[order(tab, decreasing=TRUE)][1:50]

# comment 1 when it contains "liver"
tab_liver <- tab[grepl("liver", names(tab), ignore.case=TRUE)]
# tab_liver[order(tab_liver , decreasing=TRUE)][1:50]

# Remove spaces around colon
cmnt <- cemp_t$cmnt1
cmnt2 <- gsub(" : ", ":", cmnt, fixed = TRUE)
cmnt2 <- gsub(": ", ":", cmnt2 , fixed = TRUE)
cmnt2 <- gsub(" :", ":", cmnt2 , fixed = TRUE)
cmnt2 <- gsub(" is ", ":", cmnt2 , fixed = TRUE)
cmnt2 <- gsub("lll", "ll", cmnt2 , fixed = TRUE)
cmnt2 <- gsub("hh", "h", cmnt2 , fixed = TRUE)
cmnt2 <- gsub("gellow", "yellow", cmnt2 , fixed = TRUE)
cmnt2 <- tolower(cmnt2)

containing_liver_colour <- grepl("liver colour:", cmnt2, ignore.case=TRUE) | grepl("liver colour", cmnt2, ignore.case=TRUE)
sum(containing_liver_colour)

# cmnt3 below is a "blind alley", was made only in order to print out a bunch of colur codes (below, we continue form cmnt2)
cmnt3 <- sub("liver colour:", "", cmnt2 , fixed = TRUE)
cmnt3[!containing_liver_colour] <- NA
tab_livercolour <- table(cmnt3)
tab_livercolour <- tab_livercolour[order(tab_livercolour, decreasing=TRUE)]
nm <- names(tab_livercolour)
# for printing, avoid some long ones
sel <- !substr(nm,1,4)=="part" & nchar(nm) <= 20
nm[sel]


# cmnt2 still has the same length as the data:
length(cmnt2)

str_cols <- c("yellow","brown","red","white","grey","green","pink")

#
# Make boolean matrix for colours in comment 1, with 8 rows (colors) and 30259 columns
#
containing_colour_matrix1 <- aaply(str_cols, 1, function(string) grepl(string, cmnt2, ignore.case=TRUE) & containing_liver_colour)
dim(containing_colour_matrix1)
apply(containing_colour_matrix1, 1, sum)


#
# Make boolean matrix for colours in comment 2
#
# All those with liver colour in column 1
at_least_one_color_column1 <- apply(containing_colour_matrix1, 2, sum) > 0
sum(at_least_one_color_column1 & containing_liver_colour)

# All those which probably has liver colour in column 2
sum(!at_least_one_color_column1 & containing_liver_colour)

# Column 2
cmntcol2 <- cemp_t$cmnt2
cmntcol2 <- tolower(cmntcol2)
# We are only interested in the cases where column 1 contains no colours, but contains the word "liver colour"
cmntcol2[at_least_one_color_column1 | !containing_liver_colour] <- NA

# Find colours in column 2
containing_colour_matrix2 <- aaply(str_cols, 1, function(string) grepl(string, cmntcol2, ignore.case=TRUE))
dim(containing_colour_matrix2)
apply(containing_colour_matrix2, 1, sum)

# Add colours from column 1 and column 2
containing_colour_matrix_sum <- containing_colour_matrix1 + containing_colour_matrix2
# Turn it in the right direction
containing_colour_matrix_sum <- t(containing_colour_matrix_sum)

containing_colour_matrix_sum <- data.frame(containing_colour_matrix_sum)
colnames(containing_colour_matrix_sum) <- paste("Liver", str_cols, sep="_")
str(containing_colour_matrix_sum)

# The mostcommon color combinations: yellow-red
# summarise
comb <- containing_colour_matrix_sum %>% 
  group_by(Liver_yellow, Liver_brown, Liver_red, Liver_white, Liver_grey, Liver_green, Liver_pink)
combsumm <- summarise(comb, n=n())
# sort
combsumm <- combsumm[order(combsumm$n, decreasing=TRUE),]
combsumm
 
#
# Add color variables to cemp_t 
#

cemp_t <- cbind(cemp_t, containing_colour_matrix_sum)

#
# Some more checking
#
# Check Comment 2
tab <- table(cemp_t$cmnt2)
tab[order(tab, decreasing=TRUE)][1:20]

# Check Comment 2 when comment 1 EQUALS "liver colour" and little else
nc <- nchar(names(tab_livercolour))
tab2 <- tab_livercolour[order(nc, decreasing=FALSE)][1:20]
tab2

# Check comment 3
tab <- table(cemp_t$cmnt3)
tab[order(tab, decreasing=TRUE)][1:5]
```

Make "compound" liver colour `Liver_color`
```{r }
colnames_cols <- paste("Liver", str_cols, sep="_")
colour_df <- cemp_t[, colnames_cols]
no_col <- apply(colour_df, 1, sum)
table(no_col)
#head(colour_df)
for (i in seq_along(colnames_cols)){
  colour_df[,i] <- as.character(colour_df[,i])
  colour_df[,i] <- sub(0, "", colour_df[,i])
  colour_df[,i] <- sub(1, str_cols[i], colour_df[,i])
  }
table(colour_df$Liver_yellow)
# colour_df$Liver_color <- paste(colour_df$Liver_yellow, colour_df$Liver_red, sep="-")
colour_df$Liver_color <- with(colour_df, 
                              paste(Liver_yellow, Liver_brown, Liver_red, 
                                    Liver_white, Liver_grey, Liver_green, Liver_pink, sep = "-"))
table(colour_df$Liver_color)

# Replace one or more hyphens with a single hyphen
colour_df$Liver_color <- gsub("-+", "-", colour_df$Liver_color)
# Remove hyphen at start
colour_df$Liver_color <- gsub("^-", "", colour_df$Liver_color)
# Remove hyphen at end
colour_df$Liver_color <- gsub("-$", "", colour_df$Liver_color)
table(colour_df$Liver_color)

# Add Liver_color to cemp_t
cemp_t$Liver_color <- colour_df$Liver_color
```

```{r}
cemp_t2 <- cemp_t
```

#### Merge cemp_t2 with cemp_c2 to create cemp_comb
```{r cemp_c2_to_cemp_comb_leftjoin_with_tissue3}
# Define parameters
key_pars <- c("myear", "seqno", "rlabo", "subno","tissu")
extra_pars <- c("tiswt","drywt","exlip","exlim","tiswtx","drywtx","fatwtx","gmlim", paste("Liver", str_cols, sep="_"))

# Merge
cemp_comb <- merge(cemp_c2, cemp_t2[,c(key_pars, extra_pars)], 
                   by=key_pars,
                   all.x=TRUE, all.y=FALSE)

# Check data
comb <- cemp_comb %>% group_by(myear, seqno, rlabo, subno, tissu, repno, param) %>% summarise(n=n())
xtabs(~n, comb)
```

### Comments of cemp_f
```{r cemp_c2_to_cemp_comb_leftjoin_with_fish1_comment_disease}
# Parameters
apply(!is.na(cemp_f2), 2, sum)

# comment 1
# Much Cryptocotyle lingua (svartprikksyke, en trematode)
tab <- table(cemp_f2$cmnt1)
tab[order(tab, decreasing=TRUE)][1:50]
sum(grepl("Cryptocotyle lingua", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("Cryptocotyle", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("lingua", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("Anisakis simplex", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("Anisakis", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("simplex", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("ulceration", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("ulcer", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("lymphocytic", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("lesion", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("caligiform copepod", cemp_f2$cmnt1, ignore.case=TRUE))  # sea louse = Caligidae
sum(grepl("caligiform", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("Caligus", cemp_f2$cmnt1, ignore.case=TRUE))
sum(grepl("Lepeophtheirus", cemp_f2$cmnt1, ignore.case=TRUE))

cemp_f2$Cryptocotyle_lingua <- grepl("Cryptocotyle lingua", cemp_f2$cmnt1, ignore.case=TRUE)
cemp_f2$Anisakis <- grepl("Anisakis", cemp_f2$cmnt1, ignore.case=TRUE)
cemp_f2$Ulceration <- grepl("Ulceration", cemp_f2$cmnt1, ignore.case=TRUE)
cemp_f2$Caligiform <- grepl("Caligiform", cemp_f2$cmnt1, ignore.case=TRUE)
```

```{r}
save(cemp_f2, cemp_t2, file="Milkys_Data_2016_01_18_cemp_f2_and_t2.RData")
```

#### Actual merging
Add columns to `cemp_comb` by merge with `cemp_f2` (which already contains jmpst and speci from cemp_m)
```{r cemp_c2_to_cemp_comb_leftjoin_with_fish2}
key_pars <- c("myear", "seqno", "rlabo", "subno")
extra_pars <- c("lnmin","lnmax","lnmea","lnstd",
  "wtmin","wtmax","wtmea","wtstd","sexco",
  "agmin","agmax","agmea",
  "gonwt","shlwt","agdet",
  "shlwtx","tiswtx",
  "imposex_sh","imposex_pe","imposex_stage",
  "jmpst","speci","sdate","lnmea2",
  "Cryptocotyle_lingua","Anisakis","Ulceration","Caligiform")

cemp_comb <- merge(cemp_comb, cemp_f2[c(key_pars, extra_pars)],  
                   by=key_pars,
                   all.x=TRUE, all.y=FALSE)

# str(cemp_comb)
tab <- xtabs(~myear + jmpst, cemp_comb)
dim(tab)
# stem(apply(tab, 2, sum))
tab[,apply(tab, 2, sum) > 2000]

colnames(cemp_comb )
```

## Check tiswt, tiswtx.x and tiswtx.y
tiswt and tiswtx.x come from the tissue table (cemp_t)
tiswtx.y comes from the fish table (cemp_f)
```{r cemp_comb_check_tissue_weight}
find_variable <- function(dataset, searchstring){
  colnames(dataset)[grepl(searchstring, colnames(dataset), ignore.case=TRUE)]
  }

find_variable(cemp_c, "tiswt")
find_variable(cemp_t, "tiswt")
find_variable(cemp_f, "tiswt")

tiswt_diff1 <- with(cemp_comb, tiswtx.x - tiswt)
tiswt_diff2 <- with(cemp_comb, tiswtx.y - tiswt)
tiswt_diff3 <- with(cemp_comb, tiswtx.y - tiswtx.x)

perc_difference <- function(x,y) 100*(x-y)/y
quantile(perc_difference(cemp_comb$tiswtx.x, cemp_comb$tiswt), (0:20)/20, na.rm=TRUE)
quantile(perc_difference(cemp_comb$tiswtx.y, cemp_comb$tiswt), (0:20)/20, na.rm=TRUE)
quantile(perc_difference(cemp_comb$tiswtx.y, cemp_comb$tiswtx.x), (0:20)/20, na.rm=TRUE)

# 48 cases with >1% difference (more than 5% in all cases)
sum(perc_difference(cemp_comb$tiswtx.y, cemp_comb$tiswtx.x) > 1, na.rm=TRUE)

# Check those - all are blue mussel in 1992
i <- which(perc_difference(cemp_comb$tiswtx.y, cemp_comb$tiswtx.x) > 5)
length(i)
cemp_comb[i, c("speci", "myear", "param")]

# We delete cemp_comb$tiswtx.y and renames cemp_comb$tiswtx,x
cemp_comb$tiswtx.y <- NULL
cemp_comb <- dplyr::rename(cemp_comb, tiswtx = tiswtx.x)
```

### Final check for duplicates
```{r cemp_comb_check_for_duplicates}
# Number of uniques
comb <- cemp_comb %>% group_by(myear, seqno, rlabo, subno, tissu, repno, param) %>% summarise(n=n)
xtabs(~n, comb)
# tab
#     1 
# 318654
```

## Save data (and param_be, which contains the names of biological effect parameters)
```{r}
save(cemp_comb, file="Milkys_Data_2016_01_18_cemp_combinedtable.RData")
```

## A bunch of plots of biological variables
```{r}
xtabs(~param, cemp_comb)
ggplot(subset(cemp_comb, param=="LNMEA"), aes(x=myear, y=valsnf)) + geom_point()
ggplot(subset(cemp_comb, param=="EROD"), aes(x=myear, y=valsnf)) + geom_point()

# ggplot(cemp_f2, aes(x=myear, y=lnmea)) + geom_point() + facet_wrap(~inorb)
# ggplot(cemp_f2, aes(x=myear, y=agmea)) + geom_point() + facet_wrap(~inorb)
# ggplot(cemp_f2, aes(x=agmea, y=lnmea)) + geom_point() + facet_wrap(~inorb)
# ggplot(cemp_f2, aes(x=lnmea, y=wtmea)) + geom_point() + facet_wrap(~inorb)
# ggplot(cemp_f2, aes(x=lnmea, y=wtmea)) + geom_point() + facet_wrap(~inorb)
```

